\clearpage
\setcounter{page}{1}

\chapter*{Syllabus}
\thispagestyle{fancy}
\fancyhead[RE,LO]{Syllabus}
The General Physics Laboratory is designed as introduction to research methods.
It focuses on applications of physics relevant to the life sciences.
Students learn how to design experiments to answer physical question, learn how to summarize and present their methods, findings and conclusions, and how to present their conclusions both in written and oral form.
Students also learn how to discuss their findings and be able to defend their conclusions.

\section*{Instructor {\small (see lab schedule)}}
Name: $\rule{10cm}{0.15mm}$ \\
\medskip \\ 
Email: $\rule{10cm}{0.15mm}$ \\
\medskip \\
Office: $\rule{10cm}{0.15mm}$ \\
\medskip \\
Office hours: $\rule{10cm}{0.15mm}$

\section*{Learning Outcomes}
By completing this lab course, students will be able to
\begin{itemize}
\item Design experiments, and measure \& analyze data to answer physical questions in the life sciences and related areas.
\item Use modern experimental equipment and computer analysis to measure various physical phenomena.
\item Present and discuss methods, findings and interpretation of data both in written and oral form.
\item Collaborate on a research project in a team.
\end{itemize}

\newpage

\section*{How to be successful in this course}
The key to being successful in this course is to focus on teamwork.
By working together to efficiently gather and analyze the data, your team will be free to spend more time discussing your results.
This is the true focus of this lab class: using your data to draw appropriate conclusions about the phenomena that you are investigating.
%So how can you make the best use of your time to get a good grade in this class?
\begin{itemize}
	\setlength\itemsep{2pt}
	\item \textbf{Before you come to lab:} Read over the entire lab.
	Spend some time thinking about what is involved and how you might do the experiment.
	Read through any relevant sections in the textbook even if the lecture class hasn't reached them yet.
	Communicate with your lab team to begin planning out the way you will carry out the experiment.
	\item \textbf{While you are in the lab:} Focus on your goals and tasks in the lab, don't waste time.
	If you take a lot of time on irrelevancies, you may have trouble finishing.
	Remember to document what you are doing in a lab report created as you go.
	\item \textbf{Before you leave the lab:} Be sure each team member has a copy of the data.
	You don't want to arrive in the second (or third) week and find the only person with your data has dropped the course (or is sick, or is away at a sports event, or forgot it, or ...).
	At the end of a one-week lab or the last week of a multi-week lab, hand in your finished lab report before you leave.
\end{itemize}

\section*{Roles}
In order to facilitate the preparation of the lab report, you will be working in groups of two or three.  There are three roles that your group members will fill; while each member takes primary responsibility for one role and for the portion of the lab report related to that role, please keep in mind that the experiment is a group effort and you should all be aware of the dilemmas faced by your peers and the decisions that they make.  Also, except when writing the report, these lab experiments often involve ``all hands on deck'' -- with every group member contributing to the construction, execution, and analysis of an experimental protocol.  The division of labor will be as follows:
\begin{list}{}{\itemsep=1pt} %\topsep=2pt
\item \textbf{The Journalist:} This person is primarily responsible for taking notes of everything that happens during the experiment and writing up the ``Introduction'' section of the lab report.
 
\item \textbf{The Data Interpreter:} This person primarily deals with tabulating and displaying the data, operating the computer, and writing up the ``Graphs \& Analysis'' section of the lab report. While all members of the group are expected to be involved in the collection and analysis of the results, this person is responsible for making them presentable.
 
\item \textbf{The Checker:} This person is responsible for making sure that the group is properly following the experiment plan, and makes sure that all requirements from the lab manual are being met. They are in charge of writing the ``Conclusion'' section. This person also acts as a ``manager'' of the lab tasks, stepping in where help is needed and coordinating the group's efforts to ensure the lab is completed efficiently and on-time.
\end{list}

\newpage

\section*{Schedule}
The table below shows the typical schedules for Physics 2131.
This schedule is subject to change due to unforeseen circumstances, any changes will be announced via blackboard.
\begin{table}[ht]
	\centering
	\begin{tabular}{ |c|c|c| } 
	 \hline
	 \textbf{Week} & \textbf{Fall \& Winter Semesters} & \textbf{Spring/Summer Semester} \\ 
	 \hline
	 1 & Excel Introduction & Excel Introduction \\ 
	 \hline 
	 2 & \multirow{2}{*}{Experiment 1} & \multirow{2}{*}{Experiment 1} \\ 
	 \cline{1-1}
	 3 &  &  \\ 
	 \hline
	 4 & \multirow{2}{*}{Experiment 2} & \multirow{2}{*}{Experiment 2} \\ 
	 \cline{1-1}
	 5 &  &  \\ 
	 \hline
	 6 & \multirow{3}{*}{Experiment 3} & \multirow{2}{*}{Experiment 3} \\ 
	 \cline{1-1}
	 7 & &  \\ 
	 \cline{1-1} \cline{3-3}
	 8 &  & \multirow{2}{*}{Experiment 4} \\ 
	 \cline{1-2}
	 9 & \multirow{2}{*}{Experiment 4} &  \\ 
	 \cline{1-1} \cline{3-3}
	 10 &  & \multirow{2}{*}{Experiment 5} \\ 
	 \cline{1-2}
	 11 & \multirow{2}{*}{Experiment 5} &  \\ 
	 \cline{1-1} \cline{3-3}
	 12 &  & Final Presentations \\ 
	 \hline
	 13 & Final Presentations & --- --- --- --- --- --- --- --- \\ 
	 \hline
	\end{tabular}
\end{table}

\section*{Late/Absentee Policy}
If you anticipate missing a lab session, try to arrange ahead of time to attend another lab section for that session or for the entire lab unit.
If it is not possible to attend a different lab session, contact your TA as soon as you are aware of your impending absence.
Only those with a \textbf{VALID WRITTEN EXCUSE} for missing a lab will be allowed to do a makeup activity at the end of the semester (that will take at least two hours and may involve doing another lab or evaluating data).
If you do not have a valid written excuse, you will get a zero for the week that you missed.
You may make up a maximum of one excused absence.
\par
\textbf{Each unexcused absence will remove a proportional amount from your lab grade, i.e., for a two week experiment, missing one of the days means that your grade for that experiment will be reduced by half.
If you miss more than one week, you may receive an incomplete or a failing grade for the entire class.}

\newpage

\section*{Grading}
\begin{itemize}
\itemsep0em
\item \textbf{Leading a discussion:} Every student will lead a discussion 1-2 times per semester. The student will introduce a question, problem, or idea that they encountered during class that day, and lead a short discussion about it. Total time 5-7 minutes. \newline – \textbf{10\%} of grade (Individual)
% \item \textbf{Report Introduction:} Before the first day of an experiment, each student will write a draft of the introduction to the lab report. Focus on describing what you are investigating, the data that you will gather, and how you plan on analyzing it. At the beginning of class, discuss what you wrote with your group to plan out your course of experiments. - \textbf{20\%} of grade (Individual)
\item \textbf{Quiz:}  Before the first day of an experiment, each student will complete a online quiz about the upcoming experiment. The quiz is due the night before the class, and will contain question covering lab techniques, units, meaning of results, etc. \newline - \textbf{20\%} of grade (Individual)
\item \textbf{Lab reports}(see below)\textbf{:} For the final approx.\ 30 minutes of the last day of an experiment, your lab team will work on finalizing a write-up of the experiment that you performed following the format listed below. \newline - \textbf{40\%} of grade (Group)
\item \textbf{Final presentations:} Give complete overview of one of the experiments, with methods, data, discussion, questions, ideas, conclusions – 10 minutes each group \newline - \textbf{30\%} of grade (Group)
\end{itemize}

\subsection*{Grading Scale}
\begin{table}[ht]
\centering
\begin{tabular}{|c|c|}
\hline 
\rule[-1ex]{0pt}{2.5ex} Percent & Letter Grade \\ 
\hline 
\rule[-1ex]{0pt}{2.5ex} 85 - 100 & A \\ 
\hline 
\rule[-1ex]{0pt}{2.5ex} 75 - 85 & B \\ 
\hline 
\rule[-1ex]{0pt}{2.5ex} 65 - 75 & C \\ 
\hline 
\rule[-1ex]{0pt}{2.5ex} 55 - 65 & D \\ 
\hline 
\rule[-1ex]{0pt}{2.5ex} 0 - 55 & F \\ 
\hline 
\end{tabular}
\end{table}

\newpage

\section*{Lab Reports}
Each group will work together to submit one lab report for each experiment they complete.
The lab report must be submitted electronically to your lab TA at the end of each experiment, late submissions will not be accepted unless under extenuating circumstances.
It is recommended that students not actively engaged in recording or analyzing data should spend time working on other areas of the report.

\subsection*{Introduction \& Methods}
This section should begin with a description of the process you are investigating.
Describe what data you plan to gather and explain how it will be useful in your investigation.
Briefly describe the steps you took to gather that data and the methods you used to analyze your results. \\
\textbf{Requirements}
\begin{list}{-}{\topsep=0pt \itemsep=0pt}
	\item At least one paragraph in length
	\item Simple figures may be used to illustrate the experimental setup
	\item Explain any equations that you will be using in the course of your data analysis
	\item Be detailed but avoid minutia! No one needs to read the full details of every single step you took. 
\end{list}

\subsection*{Graphs \& Analysis}
This section should include graphs of the data that you mentioned in the introduction.
The lab manual will guide you on the number of graphs that you should aim to include, but feel free to add more if you feel they are relevant.
After each graph, include a few sentences explaining it: point out important features, comment on the fit of the data against any theoretical curves, etc. \\
\textbf{Requirements}
\begin{list}{-}{\topsep=0pt \itemsep=0pt}
	\item All axes must be well labeled at a legible size and with appropriate units
	\item Use reasonable significant figures
	\item Calculate uncertainty whenever possible and use error bars.
	\item Don't skimp out on the explanations after each graph, as this is probably the most critical part of any report. Don't rely on the reader to make the connection between what is in the graph and the phenomena you are investigating, tell them explicitly.
\end{list}

\subsection*{Conclusion}
In this section you want to demonstrate how all your results connect to elucidate the phenomena you set out to investigate.
Discuss the strongest and weakest areas of your investigation, and describe changes that you would make to your experimental plan if you were to repeat your work.
If possible, discuss how the things that you learned about in each experiment would be applicable to topics studied in other science classes.

\newpage

\section*{Lab Report Rubric}
%Lab reports will be handed in at the end of class on the last day of an experiment. 
%It is recommended that students not actively engaged in recording or analyzing data should spend time working on other areas of the report. 
All of the categories in the rubric below will be graded using the following scale:
\begin{itemize}
\itemsep-0.5em
\item \textbf{0} points – \textbf{Missing}
\item \textbf{1} point – \textbf{Major Revisions} 
\item \textbf{2} points – \textbf{Minor Revisions} 
\item \textbf{3} points – \textbf{No Changes Needed}
\end{itemize}

\begin{table}[h]
	\centering
	\begin{tabular}{|p{13cm}|c|}
	\hline
	\textbf{Introduction \& Methods} & \textbf{Points} \\
	\hline
	\emph{Experiment Design:} Did your group do a careful job designing the experiment; will your investigation adequately explore the topic of the lab? &  \\
	\hline
	\emph{Explanation Clarity:} Did your team explain the experiment so that it could be reproduced? &  \\
	\hline
	\textbf{Graphs \& Analysis} &  \\
	\hline
	\emph{Completeness:} Do you have the relavent graphs/tables, and do your graphs/tables seem to have enough data to properly explore the topic of the lab?  &  \\
	\hline
	\emph{Graph/table Design:} Do you graphs/tables look professional and are they easy to understand? &  \\
	\hline
	\emph{Graph/table Context:} Do you explain what the graphs/tables show and how they factor into your investigation? &  \\
	\hline
	\textbf{Conclusion} &  \\
	\hline
	\emph{Persuasiveness:} Are the conclusions that you draw supported by your data? (\textbf{x2}) &  \\
	\hline
	\emph{Critique:} Do you propose constructive changes that would improve your experiment, or do you propose ways to extend your investigation? &  \\
	\hline
	\textbf{Bonus} &  \\
	\hline
	\emph{Error Calculations:} Do you include calculations of error derived from your measurements; are your errors displayed in your graphs? (\textbf{x2}) &  \\
	\hline
	\textbf{Total} (Max 24 points + 6 bonus points) &  \\
	\hline
	\end{tabular}
\end{table}

\newpage

\section*{Students with Disabilities}
If you have a documented disability that requires accommodations, you will need to register with Student Disability Services for coordination of your academic accommodations. 
The Student Disability Services (SDS) office is located at 1600 David Adamany Undergraduate Library in the Student Academic Success Services department. 
SDS telephone number is 313-577-1851 or 313-577-3365 (TTD only). 
Once you have your accommodations in place, I will be glad to meet with you privately during my office hours or at another agreed upon time to discuss your needs. 
Student Disability Services' mission is to assist the university in creating an accessible community where students with disabilities have an equal opportunity to fully participate in their educational experience at Wayne State University.
\par
Please be aware that a delay in getting SDS accommodation letters for the current semester may hinder the availability or facilitation of those accommodations in a timely manner.
Therefore, it is in your best interest to get your accommodation letters as early in the semester as possible.

\section*{Religious Observance Policy}
Because of the extraordinary variety of religious affiliations represented in the University student body and staff, the Wayne State University calendar makes no provision for religious holidays.
It is University policy, however, to respect the faith and religious obligations of the individual.
Students who find that their classes or examinations involve conflicts with their religious observances are expected to notify their instructors well in advance so that alternative arrangements as suitable as possible may be worked out.

%The lab grade makes up part of your total course grade.
%This grade will be based on your team's lab reports and your individual participation in the lab and the class discussions.
%\emph{Your grade will not depend on whether or not your numerical results agree with some accepted standard but on how well you conceived and carried out the experiment.} \\
%
%\begin{tabular}{|p{13cm}|c|}
%\hline
%\textbf{Team Lab Report} & \textbf{Weight} \\
%\hline
%\emph{Design and thoughtfulness:} Did your team do a careful and thoughtful job in creating your experiment, and was this thought reflected in the journal? & 25\% \\
%\hline
%\emph{Clarity and completeness:} Did your team explain your experiment so that someone could reproduce it? & 25\% \\
%\hline
%\emph{Persuasiveness:} What conclusions did your team draw from your data and were you able to back up these conclusions with this data in a convincing way? & 25\% \\
%\hline
%\emph{Evaluation:} After observing the experiments of other groups, were you able to critique your own lab, propose constructive changes, or explain why your experiment was better than those of your classmates?  (The question you are answering in your evaluation is, ``If I got to re-do this experiment next week, how would I do it differently?'') & 25\% \\
%\hline
%\textbf{Individual Participation} & \\
%\hline
%\emph{Contribution to team presentation:} Did you participate constructively in your own group's work (protocol development and data collection)?  Did you actively participate in both the preparation of the report/presentation and its delivery? & 10\% \\
%\hline
%\emph{Contribution of other teams' presentations:} Did you ask useful questions or make comments that were valuable to the other teams' reports of their evaluations?  Did you participate in both class and small group discussion? & 10\% \\
%\hline
%\end{tabular}
%
%\begin{table}[h]
%	\centering
%	\begin{tabular}{ |c|c|l|c| } 
%	 \hline
%	 \textbf{Week} & \textbf{Week Begins} & \textbf{Content}  & \textbf{Intro Due.} \\ 
%	 \hline
%	    & Jan. 9 & \textbf{No Class} & \\ 
%	 \hline 
%	%  1 & Jan. 9 & Introduction \& Equipment walkthrough \\ 
%	% \hline 
%	    & Jan. 16 & \textbf{No Class} & \\ 
%	 \hline
%	 1 & Jan. 23 & \multirow{2}{*}{Experiment 1} & \\ 
%	 \cline{1-2} \cline{4-4}
%	 2 & Jan. 30 & & \\ 
%	 \hline
%	 3 & Feb. 6 & \multirow{2}{*}{Experiment 2} &  \\ 
%	 \cline{1-2} \cline{4-4}
%	 4 & Feb. 13 & & \\ 
%	 \hline
%	 5 & Feb. 20 & \multirow{3}{*}{Experiment 3} & Yes \\ 
%	 \cline{1-2} \cline{4-4}
%	 6 & Feb. 27 & & \\ 
%	 \cline{1-2} \cline{4-4}
%	 7 & March 6 & & \\ 
%	 \hline
%	   & March 13 & \textbf{Spring Break (No Class)} & \\ 
%	 \hline
%	 8 & March 20 & \multirow{2}{*}{Experiment 4} & Yes \\ 
%	 \cline{1-2} \cline{4-4}
%	 9 & March 27 & & \\ 
%	 \hline
%	 10 & April 3 & \multirow{2}{*}{Experiment 5} & Yes \\ 
%	 \cline{1-2} \cline{4-4}
%	 11 & April 10 & & \\ 
%	 \hline
%	 12 & April 17 & Final Presentations & \\ 
%	 \hline
%	\end{tabular}
%\end{table}